Structured Versus Decorated Cospans

### 2022-03-12 ResponseToReviews.tex

{\bf We have moved the natural isomorphism triangle as equation (1) in the statement of Theorem 2.1, and we have referred to it at page 10.} {\chris 
$\checkmark$} {\kenny $\checkmark$}

\item p. 9, ingredients for constructing a double functor of decorated cospan double categories.

Does the natural transformation $\theta \colon EF \Rightarrow F’ H$ have to be a natural isomorphism? Quite possibly I missed something, but I don’t see where in 
the following proofs that assumption is used.

{\bf We thank the referee for catching this important mistake!   It does not need to be an isomorphism, so we changed the construction in p. 
9 accordingly. In fact, in Section 6 we construct a double functor
\[      \graysquare \colon F \lCsp \to D \lCsp \]
sending any open Petri net with rates to its corresponding open dynamical system, and we do this
using a monoidal natural transformation $\theta$ that is \emph{not} an isomorphism.  We had mistakenly written that it was; we have fixed this now.}
{\john $\checkmark$} {\kenny $\checkmark$} {\chris $\checkmark$}

%{\kenny I think they are correct in that we don't use this extra assumption.}

%{\chris I have a vague feeling that we discussed something similar a long time ago, but I do not remember the context at all. Perhaps it was an 
%identity then? In any case, if we don't actually use it then we should make it into a transformation.}

%{\kenny Yes, originally it was an identity. Structured cospans has similar squares for its maps, but those ones are required to be invertible. In my 
%thesis, I was going to try to come up with a recipe for turning those squares into these ones, so I made these ones invertible, too, but never 
%actually got around to this.}

%{\chris I see! So for maps of structured cospans, invertibility is actually essential?}

%{\kenny Yep, we needed certain maps $L' F \Rightarrow F' L$ (using structured cospan letters) to be invertible. We also built those maps by hand, 
%though, and here, we're using Shulman's mystical machine to build these ones. I'm not sure what happens or is required when building maps of 
%structured cospans using Shulman's machine...} {\chris Interesting!}

@@

{\john  As Christina pointed out on Zulip, the $\theta : EF \Rightarrow F'H$ on page 26, where we are
constructing a map from open Petri nets with rates to open dynamical systems, is \emph{not} an
isomorphism.  I changed ``natural isomorphism'' to ``natural transformation'' at three points on page 26 --- 
{\bf please help make sure we catch them all!}

Also, I'd better change the remark about  $\theta: EF \Rightarrow F'H$  on page 9, and
 in the statement of Theorem 2.5 --- these are two cases where we called it a ``natural 
isomorphism''.    I'll do this now: I'll change it to a ``natural transformation''.

I've thanked the referee profusely for catching this problem --- check out my response in boldface
above, and if you're happy, we can check this one off!}

{\chris The response looks great, see my small addition above `` so we changed the construction in p. 
9 accordingly''. I seached ``isomorphism'' the whole document and I think we are fine (but perhaps 
Kenny can double check). Except this one place where I changed it to transformation above Theorem 2.5, check that it doesn't break anything? It looks 
OK to me.}

@@

%{\chris Ah you are right. I don't mind, we could specify that for function composition we sometimes use concatenation!}

%{\john There were a couple places where I put in $\circ$ --- that's what the referee is really complaining
%about here.  Almost always we don't.  I could remove the $\circ$s.  There was a reason I put them in: there are some places where writing a 
%composite 
%using plain old concatenation looks weird to me.   But I don't think it's that big a deal.}

%{\chris By searching our paper, I find $\circ$ in many places that should stay in my opinion, for composition of things that are not functions 
%necessarily, and maybe only a couple for functions namely towards the end of 6.4. I also don't think it's a big deal.}

%{\chris I made the above edit, see if you agree or change at will.}

%{\kenny I think we use concatenation far more than $\circ$, right? I'm gonna' slightly change it to ``concatenation or occasionally $\circ$'' both in 
%the response comment above and the paper. Otherwise, looks good.}

%{\john I changed ``composition of maps'' to ``composition of morphisms'' above because people often use ``map'' to mean a morphism in a category of 
%sets with extra structure, and here we just mean composition of morphisms, or sometimes horizontal 1-cells, which I hope is close enough.  I hope 
%that's okay!}

%{\chris Sure, thanks to both!}

@@

\item p3: “Using the Grothendieck construction” - maybe worth mentioning this is the “covariant” version which is different to the most common 
convention 
(at least I think so, but the literature seems to be very split over this)

{\bf It is indeed the case that we use what some people call ``covariant'' version of the Grothen\-dieck construction in this paper, however we 
believe 
that mentioning this in the introduction could put people off if they haven't heard of it. We added a comment about this in the 
appendix.} {\chris $\checkmark$} {\kenny $\checkmark$} {\john $\checkmark$}

@@

\iffalse
{\chris Not sure that `covariant' would make this clearer...both are very well the Grothendieck construction I would think, but I don't mind that 
much.}

{\kenny Agree with Christina.}

{\chris See if you like the answer.}
\fi

%{\kenny Looks good.}

%{\john Great.  Except for one thing: I hate footnotes.   If something is worth saying, just say it, don't stick a little number in the text and make 
%the reader decide if it's worth looking down at the bottom of the page to see what this mysterious number means.  That's more distracting than just 
%saying it!   So I moved the comment into the text and merged it into the previous sentence.}

%{\chris OK!}

@@

%{\john It's hard to tweak the statement, since we're really trying to introduce the notation of decorations in maps of structured cospans here!    
%So 
%I've added something about this \emph{after} the statement of Theorem 2.1.  See if you like it:
%\begin{quote} 
%Note that the decoration $s \in F(m)$ is now an object in the category $F(m)$, not an element of a set as in Fong's original approach. 
%\end{quote}}

%{\chris Looks good!}

\item Proof of theorem 2.1: You should probably at least handwave that the cartesian liftings really are cartesian liftings. (No doubt it’s easy, but 
this 
is a classic sort of spot where errors can hide.)

\textbf{Kenny Courser carefully checked that these really are co/cartesian liftings, but the argument is straightforward uninteresting so we will just 
add remarks saying that this can be shown.}   {\john $\checkmark$} {\chris $\checkmark$}

@@

{\kenny I think we originally had these details in our proof (of both Theorem 2.1 and Theorem 3.1) but removed them.}

{\john It's your evil advisor who removed those details.   I'll ``handwave it''---see above.}

\item On page 9 during the construction of $\mathbb{H}$, I would appreciate a sentence giving some intuition for the functor $E$, at this point I’m confused about  its  role.

{\bf  We have added a remark on page 9 discussing this.}  {\john $\checkmark$} {\chris $\checkmark$}

%{\color{blue}{Kenny: We don't have any examples where $E$ is not an identity, but in my mind it's a functor that possibly modifies decorations, e.g. 
%$F(a)$ is a category of graph structures on $a$, and $E(F(a))$ could be something like a category of possibly labeled graph structures on $a$, if 
%that 
%makes sense. Something potentially different than \emph{just} a graph structure on $a$.}}

%{\chris Right! We can perhaps add a sentence to this end, saying that although we only use E as an identity, someone could need to change 
%decorations 
%for some application?}

%{\kenny Sure, maybe sometime before the official statement of the Theorem when we're describing how the double functor $\mathbb{H}$ is defined?}

%{\chris Sounds good to me!}

### 2022-03-12 ResponseToReviews.tex

\section{Our Answer}

\begin{enumerate}

\item As the only point against, I feel there is a growing number of “cospan” articles. The present work seems like it could serve as a comparison of 
all of 
these, but perhaps that comparison could be more explicit and detailed: what is it done in each one of these “cospan” articles? is there anything 
that 
is not already explained in this one? If some of the articles about decorated cospans were found to contain errors, is this the only construction of 
decorated cospans that solves these errors?

{\bf The present work reviews all previous work on decorated and structured cospans in Section 6.  As we mention, our previous paper ``Structured 
cospans'' fixed the errors in previous work using decorated cospans by using structured cospans instead.   Here we describe how the new improved 
decorated cospans can also fix those errors.     There may be even more ways to correct these errors, but we do not know them.}  {\kenny $\checkmark$}

@@

{\bf Done.}

{\chris $\checkmark$} {\john $\checkmark$}  {\kenny $\checkmark$}

\item A last general comment is that the paper unfortunately adopts the strategy, all too common in math papers, of putting all the examples in a 
section at 
the end of the paper. I think the readability of the paper would be improved, without much change to the organization, if the motivating example of 
open graphs were added to the relevant places in Sections 2 and 3. The remaining examples could be left in Section 6.

{\bf We believe that it is hard to analyze the examples (even the motivating one of open graphs) without having laid all relevant theory. Also, 
the comparison between the two descriptions -- prominent in Section 6.1 -- would then not be possible to illustrate. In order to 
better direct the reader to the applications, we added one sentence at each introductory paragraph of Sections 2 and 3 that point to Section 6 for 
concrete examples.} {\chris $\checkmark$} {\kenny $\checkmark$}

@@

%{\john It's hard to tweak the statement, since we're really trying to introduce the notation of decorations in maps of structured cospans here!    
%So 
%I've added something about this \emph{after} the statement of Theorem 2.1.  See if you like it:
%\begin{quote} 
%Note that the decoration $s \in F(m)$ is now an object in the category $F(m)$, not an element of a set as in Fong's original approach. 
%\end{quote}}

%{\chris Looks good!}

\item Proof of theorem 2.1: You should probably at least handwave that the cartesian liftings really are cartesian liftings. (No doubt it’s easy, but 
this 
is a classic sort of spot where errors can hide.)

\textbf{Kenny Courser carefully checked that these really are co/cartesian liftings, but the argument is straightforward uninteresting so we will just 
add remarks saying that this can be shown.}   {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

@@

{\kenny I think we originally had these details in our proof (of both Theorem 2.1 and Theorem 3.1) but removed them.}

{\john It's your evil advisor who removed those details.   I'll ``handwave it''---see above.}

\item On page 9 during the construction of $\mathbb{H}$, I would appreciate a sentence giving some intuition for the functor $E$, at this point I’m confused about  its  role.

{\bf  We have added a remark on page 9 discussing this.}  {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

### 2022-03-15 ResponseToReviews.tex

\item Section 6: as a reader, I really appreciate the examples: are there more in the literature? I do not think it is necessary to go in detail over each one of them, but perhaps it would be useful to point to other places where they are described.

{\bf We have added a word about Markov processes in Section 6.2, and with this, all the applications of structured or decorated cospans that we know 
of are discussed here.} {\john $\checkmark$} {\chris $\checkmark$} {\kenny $\checkmark$}

\vskip 1em

{\chris Second review starts here}
\item Given the title and abstract of this paper “Structured vs Decorated Cospans” I would expect this paper to provide a judgment that structured  cospans are strictly better than decorated cospans in all circumstances and a justification for that claim, or a detailed analysis of the pros and cons of 
structured vs decorated cospans, some case studies in applications that show how the difference between these double categories is felt in the 
application domain, and a some description of the tradeoffs. This paper should come with some practical advice for the applied category theorist on 
how to choose whether to model their application with structured vs decorated cospans.

{\bf In fact our goal, as explained in the abstract and introduction, was to prove that the structured and decorated cospan approaches are equivalent under certain conditions, and to illustrate this result with a wide variety of examples.  These examples can serve as a guide for other similar applications, while new applications may require exploring which approach works best.  We do not make a stand that either is strictly better.   However, we have changed the end of the Introduction to read as follows:

\begin{quote}
This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section 6 with applications to graphs, electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we apparently need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.
\end{quote}
}

{\john $\checkmark$} 
\iffalse
{\kenny I think the reader should come to their own conclusion about which framework they should use if they are trying to decide between the two, assuming both are applicable. I wouldn't say that one is better than the other. Personally, I would rather try structured cospans first, because I would rather think about a left adjoint than a symmetric lax monoidal pseudofunctor, but other people might feel differently.}

@@

{\chris I agree with both, and also I like the boldface above. I could add ``Those examples should serve as a guide for other similar applications, 
whereas new applications would require exploring which approach works best, for example if a left adjoint (to start off with structured cospans) or a 
lax monoidal pseudofunctor (to start off with decorated cospans) is available. We do not make a stand that either is strictly better.'' In fact, we 
could add something soft along those lines somewhere in the introduction to satisfy the reviewer, if we like.}

{\chris By the way, there is this sentence ''However, it should be clear by now that so
far, in cases where either structured or decorated cospans can be used, structured cospans are simpler. '' in section 6. Perhaps we could even imply 
that somewhere in the intro as well?}

{\kenny The boldface looks good to me, but I also like Christina's suggestions above if we really want to make an effort to appease this referee. I'm fine either way on this.}

{\john See if you like my new remarks in boldface, and give me a check mark if you do!  Kenny had checked this one off already, but I've changed it a lot.}

### 2022-03-15 cospans_new.tex
dealt with comment 9

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/03/11 - dealing with referee's comments

@@

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this with applications to electrical circuits, Petri nets, dynamical systems, and epidemiological modeling.  In particular, we describe a map from open Petri nets with rates to open dynamical systems that takes advantage of \emph{both} structured and decorated cospans.

>>>>

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/03/15 - dealing with referee's comments

@@

This result shows that under certain conditions, structured and decorated cospans provide equivalent ways of describing open systems.  We illustrate this in Section \ref{Applications} with applications to electrical circuits, Markov processes, Petri nets, Petri nets with rates, and dynamical systems.   This is meant to be a fairly thorough review of the existing literature.   It becomes clear that when either structured or decorated cospans can be used, structured cospans are simpler.  However, for open dynamical systems we apparently need decorated cospans, for reasons we explain.   Thus, to describe a map from open Petri nets with rates to open dynamical systems, we use \emph{both} structured and decorated cospans.
